Formalizing Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}

### 2022-12-14 nakayama/paper.tex
Offering a few remarks on the proof code

Collecting the \(u_{1}\) terms on the left-hand side of the equation, applying
the existence of additive inverses in $R$, and distributivity of scalar
multiplication for modules, we see that
\begin{equation}\label{equation1: e}
(1 - a_{1}) \cdot u_{1} =
  a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}\text{.}
\end{equation}
In Coq, this bit looks like the following code below.
\begin{minted}[linenos,firstnumber=157,stepnumber=5]{coq}
  assert ((Rone [+] Rminus x1) <.> u1 =M=
  linear_combin Madd Mzero action coeffs' generatingSet').
  { setoid_rewrite (module_distrib_Radd Radd Rmul Rone
      Mequiv Madd Mzero Mminus action).
    setoid_rewrite (module_Rone Radd Rmul Rone
      Mequiv Madd Mzero Mminus action).
    setoid_rewrite (module_minus_l Requiv Radd Rzero Rminus Rmul Rone
      Mequiv Madd Mzero Mminus action).
    symmetry.
    apply (group_move_r Mequiv Madd Mzero Mminus).
    symmetry.
    assumption. }
\end{minted}

Since a proper ideal contains no units and $a_1$ is taken to be in
$\mathfrak{m}$, it is not a unit. Hence, by the formalized statement mentioned
earlier, $1 - a_{1}$ is a unit of $R$. Let $b_{1} \in R$ denote its 
multiplicative inverse, so that $b_1 \cdot  (1 - a_{1}) = 1$. Multiplying this
element on the either side of \eqref{equation1: e}, we deduce the following.  
\begin{align*}
b_1 \cdot \left((1 - a_{1}) \cdot u_{1} \right)
  & = b_1 \cdot (a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}) \\
\left( b_1 \cdot (1 - a_{1})\right) \cdot u_{1}
  & = b_1 \cdot (a_{2} \cdot u_{2}) + \cdots + b_1 \cdot (a_{n+1} \cdot u_{n+1}) \\
u_1 = 1 \cdot u_{1}
  & = (b_1 \cdot  a_{2}) \cdot u_{2} + \cdots + (b_1  \cdot  a_{n+1}) \cdot u_{n+1}  
\end{align*}
\begin{minted}[linenos,firstnumber=181,stepnumber=5]{coq}
  pose proof
    (local_comm_ring_sub_1_nonunit Requiv Radd Rzero Rminus Rmul Rone
      R_local x1 Hx1_nonunit) as H1mx1_unit.
  (* u1 = (b1 coeffs') . generatingSet *)
  inversion_clear H1mx1_unit as [b1 Hb1].
  setoid_rewrite (commutative Requiv Rmul) in Hb1.
  apply (module_op_l Requiv Mequiv action) with (r:=b1) in H.
  setoid_rewrite <- (module_distrib_Rmul Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H.
  setoid_rewrite Hb1 in H.
  setoid_rewrite (module_Rone Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H.
  setoid_rewrite <- (module_linear_combin_mul_l Requiv Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H.
\end{minted}
In particular, the generator $u_1$ can be written as a scalar combination of
the $n$ generators $u_2, \ldots, u_{n+1}$, and is superfluous. Hence $M$ can be
generated by $n$ elements, and $M=0$ by the inductive hypothesis.

>>>>

Collecting the \(u_{1}\) terms on the left-hand side of the equation, applying
the existence of additive inverses in $R$, and distributivity of scalar
multiplication for modules, we see that
\begin{equation}\label{equation1: e}
(1 - a_{1}) \cdot u_{1} =
  a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}\text{.}
\end{equation}
In Coq, we worked backwards to establish this fact. The comments following
each tactic offers a brief explanation of the effects on the goal.
\begin{minted}[linenos,firstnumber=157,stepnumber=5]{coq}
  assert ((Rone [+] Rminus a1) <.> u1 =M= (* (1 - a1) u1 = ... *)
    linear_combin Madd Mzero action coeffs' generatingSet').
  { setoid_rewrite (module_distrib_Radd Radd Rmul Rone
      Mequiv Madd Mzero Mminus action). (* 1 u1 + (-a1) u1 = ... *)
    setoid_rewrite (module_Rone Radd Rmul Rone
      Mequiv Madd Mzero Mminus action). (* u1 + (-a1) u1 = ... *)
    setoid_rewrite (module_minus_l Requiv Radd Rzero Rminus Rmul Rone
      Mequiv Madd Mzero Mminus action). (* u1 - a1 u1 = ... *)
    symmetry. (* ... = u1 - a1 u1 *)
    apply (group_move_r Mequiv Madd Mzero Mminus). (* ... + a1 u1 = u1 *)
    symmetry. (* u1 = ... + a1 u1 *)
    assumption. }
\end{minted}

Since a proper ideal contains no units and $a_1$ is taken to be in
$\mathfrak{m}$, it is not a unit. Hence, by the formalized statement mentioned
earlier, $1 - a_{1}$ is a unit of $R$. Let $b_{1} \in R$ denote its 
multiplicative inverse, so that $b_1 \cdot  (1 - a_{1}) = 1$. Multiplying this
element on the either side of \eqref{equation1: e}, we deduce the following.  
\begin{align*}
b_1 \cdot \left((1 - a_{1}) \cdot u_{1} \right)
  & = b_1 \cdot (a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}) \\
\left( b_1 \cdot (1 - a_{1})\right) \cdot u_{1}
  & = b_1 \cdot (a_{2} \cdot u_{2}) + \cdots + b_1 \cdot (a_{n+1} \cdot u_{n+1}) \\
u_1 = 1 \cdot u_{1}
  & = (b_1 \cdot  a_{2}) \cdot u_{2} + \cdots + (b_1  \cdot  a_{n+1}) \cdot u_{n+1}  
\end{align*}
\begin{minted}[linenos,firstnumber=181,stepnumber=5]{coq}
  pose proof
    (local_comm_ring_sub_1_nonunit Requiv Radd Rzero Rminus Rmul Rone
      R_local x1 Hx1_nonunit) as H1ma1_unit.
  (* Getting to u1 = (b1 coeffs') . generatingSet *)
  inversion_clear H1ma1_unit as [b1 Hb1]. (* 1 - a1 is a unit *)
  setoid_rewrite (commutative Requiv Rmul) in Hb1. (* b1 (1 - a1) = 1 *)
\end{minted}
This bit of proof code has the effect of naming \(b_{1}\) as the inverse of the
unit \(1 - a_{1}\) and storing in the hypothesis \texttt{Hb1} the equation
\(b_{1} (1 - a_{1}) = 1\). The remainder of the code goes on to multiply on the
left by \(b_{1}\) in the equation \texttt{H} in order to say that
\(b_{1} ((1 - a_{1}) u1) = b_{1} (a_{2} u_{2} + \cdots + a_{n+1} u_{n+1})\).
Going on to use the rules of module actions and hypothesis \texttt{Hb1}, we can
simplify the left hand side of \texttt{H} to be just \(u_{1}\).
\begin{minted}[linenos,firstnumber=last,stepnumber=5]{coq}
  apply (module_op_l Requiv Mequiv action) with (r:=b1) in H.
  setoid_rewrite <- (module_distrib_Rmul Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H. (* H: (b1 (1 - a1)) u1 = ... *)
  setoid_rewrite Hb1 in H. (* H: (1) u1 = ... *)
  setoid_rewrite (module_Rone Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H. (* H: u1 = ... *)
\end{minted}
In particular, the generator $u_1$ can be written as a scalar combination of
the $n$ generators $u_2, \ldots, u_{n+1}$, and is superfluous. Hence $M$ can be
generated by $n$ elements, and $M=0$ by the inductive hypothesis.
